Higher order unification for free!

### 2024-04-08 paper/main.tex
update tex

\usepackage{myTools}
\usepackage{macros}
\usepackage{enumitem}
\usepackage{intcalc}

@@

% This property can be relaxed, to accept approximations: a dedicated
% section is given in a future section
\begin{lemma}
  If the \lhs of a \linkbeta is instantiated to a rigid term and its \rhs
  counterpart is still in \notllambda,
  the original unification problem is not in \llambda
  % \noteE[inline]{non capisco cosa vuoi dire con "the current...". Data la prova sotto vuoi dire che il problem originale sarebbe stato fuori... e quindi falliamo}
  %\noteE[inline]{Is it clearer?}
  and the unification fails.
  \label{lemma:keep-fidelity}
\end{lemma}

@@

@@

\newcommand{\progBetaLL}{\emph{progress-beta-\llambda}}
\newcommand{\progBetaRH}{\emph{progress-beta-rigid-head}}
\newcommand{\progBetaDedup}{\emph{progress-beta-dedup}}

@@

\noteD{added the def below to respect fidelity}
% Example of this `X a = Y a'
\begin{definition}[\progBetaDedup]
  Given two \linkbeta $l1$ and $l2$, if they share the same \lhs, unification
  fails.
  \label{def:progBetaDedup}
\end{definition}

@@

\begin{proof}[Proof sketch]
  Let $l$ a \linkbeta in the store \linkStore. If $l$ is activated by
  \progBetaRH, then it disappears from \linkStore and \elpiIn{progress}
  terminates. Otherwise, the \rhs of $l$ is made by a variable applied to $m$
  arguments. At each activation of \progBetaLL, $l$ is replaced by a new
  \linkbeta $l^1$ having $m-1$ arguments. At the $m^{th}$ iteration, the
  \linkbeta $l^m$ has no more arguments and is removed from \linkStore.
  Note that at the $m^{th}$ iteration, $m$ new \linketa have been added to
  \linkStore, however, by \cref{lemma:prog-eta-terminates}, the algorithm
  terminates.
  \noteD{L'ho riscritto}

  % \Cref{def:progBetaRH} makes \elpiIn{progress} terminates, since it makes a
  % \linkbeta disappear from \linkStore. On the other hand, \cref{def:progBetaLL},
  % creates each time a new \linketa and replace the old \linkbeta with a new one.
  % This progression can however triggered at most $m$ times, since each new
  % \linkbeta as a smaller list of applied variables. At the $m^{th}$ progression,
  % the \linkbeta is removed from \linkStore which is now filled by $m$ new
  % \linketa. By \cref{lemma:prog-eta-terminates}, we know that \elpiIn{progress}
  % terminates if \linkStore is made by only \linketa and therefore
  % \elpiIn{progress} still terminates.
  %\noteE[inline]{is it ok?}
  \noteE[inline]{funziona. per essere più precisi io parlerei di ordine lessicografico (tipico ordine ben fondato usato per dimostrare terminazione). Nl nostro caso è la tripla (argomenti extra dei beta, numero di beta, numero di eta).}
\end{proof}

@@

\subsection{Relaxing \cref{lemma:keep-fidelity} : tobe renamed}

>>>>

\usepackage{myTools}
\usepackage{macros}
\usepackage{enumitem}
\usepackage{intcalc}
\usepackage{nameref}

@@

% This property can be relaxed, to accept approximations: a dedicated
% section is given in a future section
\begin{lemma}[\linkbeta with rigid \lhs]
  If the \lhs of a \linkbeta is instantiated to a rigid term and its \rhs
  counterpart is still in \notllambda,
  the original unification problem is not in \llambda
  % \noteE[inline]{non capisco cosa vuoi dire con "the current...". Data la prova sotto vuoi dire che il problem originale sarebbe stato fuori... e quindi falliamo}
  %\noteE[inline]{Is it clearer?}
  and the unification fails.
  \label{lemma:keep-fidelity}
\end{lemma}

@@

% From \cref{lemma:keep-fidelity}, we can deduce the following corollary

% \begin{corollary}
%   Given a \linkbeta \linkbetaM{\Gamma}{X}{T}, if it exists a second link
%   $l\in\linkStore$, then unification fails.   
% \end{corollary}

@@

\newcommand{\progBetaLL}{\emph{progress-beta-\llambda}\xspace}
\newcommand{\progBetaRH}{\emph{progress-beta-rigid-head}\xspace}
\newcommand{\progBetaDedup}{\emph{progress-beta-dedup}\xspace}

@@

\noteD{added the def below to respect fidelity}
% Example of this `X a = Y a'
\begin{definition}[\progBetaDedup]
  Given a \linkbeta $l_1$ and second link $l_2 \in\linkStore$, such that they
  share the same \lhs. In this case, the two \rhs are unified and a $l_2$ is
  removed from \linkStore.
  \label{def:progBetaDedup}
\end{definition}

@@

\begin{proof}[Proof sketch]
  Let $l$ a \linkbeta in the store \linkStore. If $l$ is activated by
  \progBetaRH, then it disappears from \linkStore and \elpiIn{progress}
  terminates. Otherwise, the \rhs of $l$ is made by a variable applied to $m$
  arguments. At each activation of \progBetaLL, $l$ is replaced by a new
  \linkbeta $l^1$ having $m-1$ arguments. At the $m^{th}$ iteration, the
  \linkbeta $l^m$ has no more arguments and is removed from \linkStore.
  Note that at the $m^{th}$ iteration, $m$ new \linketa have been added to
  \linkStore, however, by \cref{lemma:prog-eta-terminates}, the algorithm
  terminates. Finally \progBetaDedup also guarantees termination since it 
  makes a unification $\mathbb{U}$ and if $\mathbb{U}$ fails, then 
  \elpiIn{progress} terminates and if $\mathbb{U}$ succeeds, 
  the recursive calls to \elpiIn{progress} have a specialized \linkbeta\dots

  % \Cref{def:progBetaRH} makes \elpiIn{progress} terminates, since it makes a
  % \linkbeta disappear from \linkStore. On the other hand, \cref{def:progBetaLL},
  % creates each time a new \linketa and replace the old \linkbeta with a new one.
  % This progression can however triggered at most $m$ times, since each new
  % \linkbeta as a smaller list of applied variables. At the $m^{th}$ progression,
  % the \linkbeta is removed from \linkStore which is now filled by $m$ new
  % \linketa. By \cref{lemma:prog-eta-terminates}, we know that \elpiIn{progress}
  % terminates if \linkStore is made by only \linketa and therefore
  % \elpiIn{progress} still terminates.
  %\noteE[inline]{is it ok?}
  \noteE[inline]{funziona. per essere più precisi io parlerei di ordine lessicografico (tipico ordine ben fondato usato per dimostrare terminazione). Nl nostro caso è la tripla (argomenti extra dei beta, numero di beta, numero di eta).}
  \noteD{da finire}
\end{proof}

@@

\subsection{Relaxing lemma \nameref{lemma:keep-fidelity}}
